Differential invariant signatures for planar Lie group transformations of images

### 2021-02-10 paper/newpaper.tex
Some more intro material on signatures etc, 3D, and why we can't do completeness

There are a variety of methods that can be used to identify differential invariants, and we provide examples of three of them: tensor contraction, transvectants, and the method of moving frames. We use the resulting invariants to construct practical sets to form three-dimensional invariant signatures. These signatures are not necessarily complete: the image cannot always be reconstructed uniquely up to transformation, but they are plottable, and depend at worst on third derivatives, although more channels of information, such as colour images, can reduce the highest order to derivate needed in some cases. We demonstrate the invariant signatures for each transformation group based on a simple smooth image. A full consideration of how these signatures could be used in practice will require effective methods to numerically approximate derivatives effectively for images.
\end{abstract}

@@

The introduction of an invariant signature reduces the recognition problem to one of comparing the higher-dimensional signatures (typically the signatures need to be at least three dimensional for planar images) that encode the desired self-equivalences. In this paper we introduce the mathematical considerations necessary to construct invariant signatures to the planar Lie groups (and some infinite-dimensional, pseudo-Lie groups too) that can transform 2D images. As we shall show, the lattice structure of the set of groups influences the relationship between the sets of invariants for the different groups. We provide sample signature sets for the various groups we consider, and examples of the appearance of those signatures for transformations of a simple, smooth image, but do not consider the question of a complete practical implementation. We do not consider the classification of the signatures, for which either a standard image recognition algorithm can be used (see, e.g.,~\cite{Zhang20} for a survey), or a matching method such as~\cite{UsCurrents} can be used.

In order for the invariant signatures to be useful to identify similarity in images, there are two key considerations: they must vary continuously with respect to the image, either for all smooth images or for all except a set of sufficiently high codimension, and they must be robust with respect to occlusions, where one object obscures some part of another. Thse constraints are essential if the signature is to be used to detect similarity of images, but the first fails for the rational invariants found in the literature. Such considerations lead one naturally to consider local invariants, i.e., those defined using each point of the object as its own basepoint, together with a neighbourhood in some topology. This is sufficient to compute, e.g., derivatives of an object. For this reason we focus on differential invariants. The disadvantage of such invariants, as we shall discuss, is the need to compute several orders of spatial derivative of the image.

Ideally, an invariant signature would be complete, so that the image is exactly determined by the signature up to a transformation, and two objects share a signature if and only if they differ only by a transformation from $\mathcal{G}$. However, this often requires signatures in many dimensions, and potentially also more derivatives of the image. For many applications an incomplete signature is sufficient, and is what we consider here. See, e.g.,~\cite{UsMobius} and references therein for more discussion about the desirable properties of a set of invariants for image and curve analysis.

@@

An $m$-dimensional signature set is the image of $m$ invariants $\mathcal{I}=(I_1,I_2,\dots,I_m)$, i.e., the set $\mathcal{I}(f)(\Omega)\subset \mathbb{R}^m$. See Section~\ref{sec:T2} for a basic example.  A signature set is complete for $f$ if it determines $f$ up to transformations from $\mathcal{G}$, i.e., if $\mathcal{I}(\tilde f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies that there exists a transformation $\varphi\in \mathcal{G}$ such that $\tilde f = \varphi\cdot f$ and $\tilde \Omega = \varphi(\Omega)$. If the signature is not complete, then there will be a \emph{bad set} of images that cannot be differentiated between.

@@

%Groups we've done:
%translations, E(2), Sim(2), SA(2), A(2), PSL(2,C), PSL(3,R), Diff\_con, Diff\_vol, Diff
\subsection{Example signatures}

In this paper our aim is develop the mathematical underpinnings of
constructing differential invariant signatures for images under the action
of planar Lie groups. However, as we develop the signature for each group,
we will provide an example of the type of surface that the signature
defines. Note that these will generally be three-dimensional representation
of the planar image.

>>>>

There are a variety of methods that can be used to identify differential invariants, and we provide examples of three of them: tensor contraction, transvectants, and the method of moving frames. We use the resulting invariants to construct practical sets to form three-dimensional invariant signatures. These signatures are not necessarily complete: the image cannot always be reconstructed uniquely up to transformation, but they are plottable, and depend at worst on third derivatives, although more channels of information, such as colour images, can reduce the highest order of derivative needed in some cases. We demonstrate the invariant signatures for each transformation group based on a simple smooth image. A full consideration of how these signatures could be used in practice will require effective methods to numerically approximate derivatives effectively for images.
\end{abstract}

@@

A greyscale image can be viewed as a two-dimensional manifold embedded in a
three dimensional ambient space, with coordinates $(x, y, f(x,y))$. A
signature has coordinates $\mathcal{I}(f) = (I_1, I_2, \ldots, I_m)$ which
are functions of the derivatives of $f(x, y))$, and will therefore be a
two-dimensional manifold embedded in $\mathbb{R}^m$. Under a change of
coordinates $(x, y) \mapsto (\bar{x}, \bar{y})$ induced by a group action,
the signature of the corresponding image $\bar{f}(\bar{x}, \bar{y}) = f(x, y)$
will have the same signature.

The introduction of an invariant signature reduces the recognition problem to
one of comparing the signatures in a higher-dimensional space that encode the
desired self-equivalences. For greyscale images, it is natural to seek
signatures with three coordinates as $\mathbb{R}^3$ is the smallest space in
which the signature surfaces can reside. In this paper we introduce the
mathematical considerations necessary to construct invariant signatures to
the planar Lie groups (and some infinite-dimensional, pseudo-Lie groups too)
that can transform 2D images. As we shall show, the lattice structure of the
set of groups influences the relationship between the sets of invariants for
the different groups. We provide sample signature sets for the various groups
we consider, and examples of the appearance of those signatures for
transformations of a simple, smooth image, but do not consider the question
of a complete practical implementation. We do not consider the classification
of the signatures, for which either a standard image recognition algorithm
can be used (see, e.g.,~\cite{Zhang20} for a survey), or a matching method
such as~\cite{UsCurrents} can be used.

In order for the invariant signatures to be useful to identify similarity in images, there are two key considerations: they must vary continuously with respect to the image, either for all smooth images or for all except a set of sufficiently high codimension, and they must be robust with respect to occlusions, where one object obscures some part of another. These constraints are essential if the signature is to be used to detect similarity of images, but the first fails for the rational invariants found in the literature. Such considerations lead one naturally to consider local invariants, i.e., those defined using each point of the object as its own basepoint, together with a neighbourhood in some topology. This is sufficient to compute, e.g., derivatives of an object. For this reason we focus on differential invariants. The disadvantage of such invariants, as we shall discuss, is the need to compute several orders of spatial derivative of the image.

Ideally, an invariant signature would be complete, so that the image is exactly determined by the signature up to a transformation, and two objects share a signature if and only if they differ only by a transformation from $\mathcal{G}$. However there are some pathological scenarios that occur regularly in images, particularly that flat regions of images collapse down to single points in the signature as all of the derivatives are zero in these regions. So images that are otherwise similar apart from the size of one of these regions cannot be distinguished between. For this reason, we do not pursue complete signatures, for many applications an incomplete signature is sufficient, and is what we consider here. See, e.g.,~\cite{UsMobius} and references therein for more discussion about the desirable properties of a set of invariants for image and curve analysis.

@@

An signature set is the image of $m$ invariants
$\mathcal{I}=(I_1,I_2,\dots,I_m)$. For simplicity of notation we will write
$\mathcal{I}(f)$ with the understanding that each signature component is
actually acting on the jet space $J^d(f, \Omega, k)$. The signature
$\mathcal{I}(f) \subset \mathbb{R}^m$. See Section~\ref{sec:T2} for a basic
example. A signature set is complete for $f$ if it determines $f$ up to
transformations from $\mathcal{G}$, i.e., if $\mathcal{I}(\tilde
f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies that there exists a
transformation $\varphi\in \mathcal{G}$ such that $\tilde f = \varphi\cdot f$
and $\tilde \Omega = \varphi(\Omega)$. If the signature is not complete, then
there will be a \emph{bad set} of images that cannot be differentiated
between.

@@

%Groups we've done:
%translations, E(2), Sim(2), SA(2), A(2), PSL(2,C), PSL(3,R), Diff\_con, Diff\_vol, Diff
\subsection{Test image}
In this paper our aim is develop the mathematical underpinnings of
constructing differential invariant signatures for images under the action
of planar Lie groups. However, as we develop the signature for each group,
we will provide an example of the type of surface that the signature
defines. Note that these will generally be three-dimensional representation
of the planar image.
